Machine Learning for Predictive Auto-Tuning

### 2011-12-13 paper/draft1.tex

In this rapidly evolving landscape, the construction of \emph{general}
development tools and libraries that \emph{fully utilize} system resources
remains a daunting task. Even within specialized architectures from the
\emph{same} vendor, like NVIDIA's Graphics Processing Units (GPUs) and the
Compute Unified Device Architecture (CUDA) \citep{nickolls2008scalable,
nvidia2011cuda}, many developers default to massive amounts of manual labor
to optimize CUDA code to specific input domains. In addition, hand-tuning
rarely generalizes well to new hardware generations or different input
domains, and it can also be error-prone or far from optimal. One of the
reason is that kernels can produce staggeringly large optimization spaces
\citep{datta2008stencil}. The problem is further compounded by the fact that
these spaces can be highly discontinuous \citep{ryoo2008program}, difficult
to explore and holding quasi-optimal solutions at the edge of ``performance
cliffs'' (in part due to hard resource restrictions, like the register file
size).

>>>>

In this rapidly evolving landscape, the construction of \emph{general}
development tools and libraries that \emph{fully utilize} system resources
remains a daunting task. Even within specialized architectures from the
\emph{same} vendor, like NVIDIA's Graphics Processing Units (GPUs) and the
Compute Unified Device Architecture (CUDA) \citep{nickolls2008scalable,
nvidia2011cuda}, many developers default to massive amounts of manual labor
to optimize CUDA code to specific input domains. In addition, hand-tuning
rarely generalizes well to new hardware generations or different input
domains, and it can also be error-prone or far from optimal. One of the
reason is that kernels can produce staggeringly large optimization spaces
\citep{datta2008stencil}. The problem is further compounded by the fact that
these spaces can be highly discontinuous \citep{ryoo2008program}, difficult
to explore and quasi-optimal solutions lie at the edge of ``performance
cliffs'' induced by device-specific hard constraints (e.g. register file
size and shared memory size).
